---
title: "Assigment1"
author: "Pavel Linder, Nikita Brancatisano"
date: "11/6/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

#install.packages("scatterplot3d")
#library("scatterplot3d")
install.packages("rgl")
library("rgl")

Task 1: Maximization of a non-convex function.
## 1. Code a method f(x, y) that computes a value z, given an input tuple (x, y).
f <- function(x, y)  {
  z <- ((1 - x)^2) + (exp(1) * (y - (x^2))^2)
}

## 2. Code a method that visualizes the Rosenbrock function in 3D
x <- y <- seq(-1, 1, length= 400)
z <- outer(x, y, f)
z[is.na(z)] <- 1 # change non-defined elements to 1
persp(x, y, z, theta = 0, phi = 25, expand = 1, col = "lightblue", ticktype = "detailed")

## 3. Code a genetic algorithm, that attempts to find the global maximum of this function. 
GA <- ga(type = "real-valued", 
         fitness =  function(x) f(x[1], x[2]),
         lower = c(-1, -1), upper = c(1, 1), 
         popSize = 50, maxiter = 1000, run = 100)
plot(GA)
summary(GA)
res <- persp(x, y, z, theta = 45, phi = 20, expand = 1, col = "lightblue",
      ltheta = 120, shade = 0.75, ticktype = "detailed")
max_x = GA@solution[,1]
max_y = GA@solution[,2]
max_z = f(GA@solution[,1], GA@solution[,2])
points(trans3d(max_x, max_y, max_z, pmat=res), col="red")
lines (trans3d(x, max_y, max_z, pmat = res), col = "blue")
lines (trans3d(max_x, y, max_z, pmat = res), col = "blue")

Task 2: Genetic feature selection
## 
# Read in the dataset 
genes <- DLBCL

# Get the summary statistics of the data
summary(genes)


